---
title: "Conditional Expected Shortfall"
subtitle: "Nonparametric Estimation"
author: |
  | [Young-geun Kim](mailto:dudrms33@g.skku.edu)
  | [ygeunkim.github.io](https://ygeunkim.github.io)
institute: "2019711358, [Department of Statistics](https://stat.skku.edu/stat/index.jsp)"
date: "`r format(Sys.time(), '%d %b, %Y')`"
bibliography: "../docs/nonparam.bib"
output: 
  beamer_presentation:
    toc: yes
    slide_level: 2
    theme: "Malmoe"
    colortheme: "dolphin"
    fonttheme: "structurebold"
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(input = inputFile, encoding = encoding, output_dir = "../static/slides")
  })
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage[normalem]{ulem}
  - \usepackage[utf8]{inputenc}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{hyperref}
  - \usepackage{enumitem}
  - \usepackage[boxruled, linesnumbered]{algorithm2e}
  - \IncMargin{1.5em}
  - \newcommand{\iid}{\stackrel{iid}{\sim}}
  - \newcommand{\indep}{\stackrel{indep}{\sim}}
  - \newcommand{\hsim}{\stackrel{H_0}{\sim}}
  - \newcommand{\ind}{\perp\!\!\!\perp}
  - \newcommand{\R}{\mathbb{R}}
  - \newcommand{\B}{\boldsymbol\beta}
  - \newcommand{\hb}{\boldsymbol{\hat\beta}}
  - \newcommand{\E}{\boldsymbol\epsilon}
  - \newcommand{\defn}{\mathpunct{:}=}
  - \DeclareMathOperator*{\argmin}{argmin}
  - \DeclareMathOperator*{\argmax}{argmax}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = .618,
  fig.pos = "H"
  )
knitr::knit_hooks$set(
  document = function(x) {
    sub("\\usepackage[]{color}", "\\usepackage{xcolor}", x, fixed = TRUE)
  }
)
options(digits = 3)
options(kableExtra.latex.load_packages = FALSE)
pander::panderOptions("digits", 3)
is_latex <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "latex"
is_beamer <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "beamer"
```

```{r, message=FALSE, echo=FALSE}
# tidyverse family---------------------
library(tidyverse)
# large data
library(data.table)
# custom packages----------------------
library(rmdtool) # devtools::install_github("ygeunkim/rmdtool")
library(goodgraphic) # devtools::install_github("ygeunkim/goodgraphic")
# set seed for report -----------------
set.seed(1)
```

```{r, message=FALSE, echo=FALSE}
# devtools::install_github("ygeunkim/ceshat")
library(ceshat)
# devtools::install_github("ygeunkim/youngtool")
library(youngtool)
# GARCH
library(fGarch)
```

## Reviewed Paper

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(RefManageR)
bib <- ReadBib("../docs/nonparam.bib")
Citet(bib, "cai:2008aa")
```

```{r, echo=FALSE, results='asis'}
PrintBibliography(bib)
```


# Simulation

## AR(1)-GARCH(1, 0)

$$
\begin{cases}
  X_t = Y_{t - 1} \\
  Y_t = 0.01 + 0.62 X_t + \sigma_t \epsilon_t \\
  \sigma_t^2 = -0.15 + 0.65 \sigma_{t - 1}^2 \\
  \epsilon_t \sim N(0, 1)
\end{cases}
$$


```{r, echo=FALSE}
garch_sim <- function(n, cond, ar_mu = .01, ar = .62) {
  garch_spec <- 
    garchSpec(
      cond.dist = "norm",
      model = list(
        omega = .15, alpha = 0, beta = .65
      )
    )
  tibble(garch = garchSim(garch_spec, n = n) %>% as.numeric()) %>% 
    mutate(
      x = cond,
      y = ar_mu + ar * x + garch
    ) %>% 
    select(y) %>% # to use youngtool (experimental stage)
    pull()
}
```

## True conditional distribution

Since $\epsilon_t \sim N(0, 1)$,

$$\sigma_t \epsilon_t \sim N(0, \sigma_t)$$

$$Y_t \mid X_t \sim N \left( 0.01 + 0.62 X_t, \sigma_t^2 \right)$$

## True CES

- For each $X_t$, `pnorm(x, mean, sd)` gives the conditional cdf value.
- Inverting $S(y \mid x) = 1 - F(y \mid x)$ gives $\nu_p(x)$.

$$\nu_p(x) = S^{-1}(p \mid x)$$

- Plugging-in method gives $\mu_p(x)$.

$$\mu_p(x) = \frac{1}{p} \int_{\nu_p(x)}^\infty y f(y \mid x) dy$$

## Goal of MC Simulation

- Compute the error between the true $\mu_p(x)$ and $\hat\mu_p(x)$
- Is the estimator of @cai:2008aa good?

## Random number generation

Monte Carlo Samples:

- For fixed $x_t$
- Generate GARCH(1, 0): $(\sigma_t, \epsilon_t)$
- $X_t = Y_{t - 1}$
- AR(1): $Y_t = 0.01 + 0.62 Y_{t - 1} + \sigma_t \epsilon_t$

```{r, echo=FALSE}
cond_sim <- function(n, m, xcond) {
  mc_data(garch_sim, N = n, M = m, cond = xcond) %>% 
    .[,
      xcond := xcond] %>% 
    .[]
}
```

```{r, echo=FALSE}
x <- runif(1)
mc <- cond_sim(200, 500, x)
```

## References


